Contextual Identity: Towards Usable Privacy

### 2013-02-28 paper.tex
more rewrites

The desire for spontaneous, positive human interaction often requires sharing
personal information. Sharing information can obviously lead to loss of
privacy, but the act of sharing alone does not negate expectation of
privacy~\cite{boyd2}.  Helen Nissenbaum has long argued that privacy violations
come not from the simple sharing of personal information, but rather from
sharing that information in a way that violates social norms, or in the wrong
context~\cite{nissenbaum}.

We consider it a contextual identity violation when
multiple contextual identities are linked without the user's permission, or the
user cannot control which contextual identity is presented to a given audience.

Lesbian, gay, bisexual, and transgendered youth are a prime example of this.
Access to support is crucial for LGTB-identified youth who are at higher risk
for bullying and suicide than non-LGTB youth, and much of that support is only
accessible through the Internet in socially conservative areas~\cite{hrc}. In
2011 Bobbi Duncan and Taylor McCormick, two students at the University of Texas
at Austin, were inadvertantly outed when their choir director added them to a
Facebook group for Queer Chorus~\cite{fowler}. Even though both students had
blocked their parents from seeing any personal posts revealing their sexuality,
because Queer Chorus is a public group Facebook published notifications to
Duncan's and McCormick's friends, including their parents. Duncan later
attempted suicide~\cite{duncan}.

@@

Barth, Datta, Mitchell, and Nissenbaum provide a formal model for contextual
integrity of user data~\cite{barth}. The contextual integrity model focuses on
the flow of personal information in terms of knowledge states held by agents
acting in given roles, and works well for modeling privacy legislation such as
HIPPA and COPPA. Two drawbacks of this work are that it is incapable of
modeling information where the subject involves more than one person, and it
only takes the type of information into account rather than the content or tone
of the information.  Thus, the event ``Bob told Alice that Charlie cooked a
delicious dinner'' is out of scope of the model.  In the context of social
networks, information is rarely ever about a single person, and the tone of the
message is as important to user happiness as the content of the message.

@@

\section{Contextual identity violations}
\label{sec:examples}
\subsection{Redistributing information out of context}
Re-distributing information out of its original context often leads to
embarrassment~\cite{nissenbaum}.  In November 2007, Facebook Beacon allowed
third-party sites to publish purchases, travel bookings, movie rentals and more
to the user's activity stream.  Because of poor opt-out and lack of visibility
into what was being published, user outcry was
immediate~\cite{mccarthy,nakashima}.  In December 2007, Google Reader exposed
RSS feeds of user-marked news stories to the user's Google Talk contacts, which
includes everyone with whom a user has chatted whether they be co-workers,
supervisors, or friends. Although this feed was always public, prior to this
launch it was not discoverable, leading to a bad experience for many
users~\cite{helft}. In February 2010, Google Buzz launched, a product that
exposed the user's most frequent Gmail and Google Talk contacts and their
publicly available (through previously less discoverable) news and photos.  The
combination of exposing contacts (which could include abusive ex-husbands,
co-workers, and friends), as well as aggressively linking photos and news
items, resulted in a huge backlash~\cite{fugitivus,buzz}. In September 2012,
Facebook imported old wall posts from 2008 into the new Timeline interface.
Although wall posts were always visible from profile pages, the new Timeline
interface brought old wall posts which users used to treat as private messages,
before the advent of ``Like'' and comment buttons, to the attention of an
audience that the user never anticipated~\cite{timeline}.

\subsection{Unsatisfiable policies}
\label{sec:policies}
Some service providers have policies which preclude isolating multiple
contextual identities. For example, Facebook and Google have a ``Real Names''
policy, which requires users to register for accounts with their legal
name~\cite{fb_names,google_names}.  These policies ignore that
community-building happens in many different contexts, that users have many
legitimate reasons for presenting different identities in different contexts,
and that users don't necessarily want those identities to be linked. For
example, disallowing avatar handles as a primary identifier makes building a
gaming community difficult.  Linking a gaming identity to a professional
identity has already caused problems for 2012 Maine Senate candidate and World
of Warcraft gamer Colleen Lachowicz~\cite{maine}.

@@

\section{Research directions}
We propose new research directions based on the following questions.
\begin{itemize}
\item How do users think about identity?
\item What are users already doing to preserve contextual identities?
\item What tasks or topics do users find embarrassing?
\end{itemize}

plethora of cross posting software, too much damn syndication indicates that
many users are on multiple platforms.

\subsection{Inferring contextual identities}
Accurately detecting the user's current contextual identity, whether
heuristically or through asking the user, is paramount to helping the user
manage it.  Although a user may subconsciously be aware of their contextual
identities, expecting users to enumerate their identities and manually curate
them is probably too burdensome. Firefox has a Profile Manager, but it is so
difficult to use that it will be removed.\footnote{See
\url{https://bugzilla.mozilla.org/show\_bug.cgi?id=214675\#c53}.}

Tang et al.~show that most people do not take the time to create labels and
maintain friend lists~\cite{tang}. However, the same authors show how to use
heterogeneous networks to infer relationship types (e.g., manager-subordinate)
though the graph itself is unlabeled.

It may also be possible to infer tasks from the user's activity. For example,
the \texttt{about:profile} Firefox extension uses the Open Directory Project
and Alexa to categorize URLs that the user has visited~\cite{aboutprofile}.

>>>>

The desire for privacy balances the desire for spontaneous, positive human
interaction, which often requires sharing personal information.  The act of
sharing alone does not negate expectation of privacy~\cite{boyd2}. Helen
Nissenbaum has long argued that privacy violations come not from the simple
sharing of personal information, but rather from sharing that information in a
way that violates social norms, or in the wrong context~\cite{nissenbaum}.

Lesbian, gay, bisexual, and transgendered youth are at risk for contextual
identity violations.  Access to support is crucial for LGTB-identified youth
who are at higher risk for bullying and suicide than non-LGTB youth, and much
of that support is only accessible through the Internet in socially
conservative areas~\cite{hrc}. In 2011 Bobbi Duncan and Taylor McCormick, two
students at the University of Texas at Austin, were inadvertantly outed when
their choir director added them to a Facebook group for Queer
Chorus~\cite{fowler}. Even though both students had blocked their parents from
seeing any personal posts revealing their sexuality, because Queer Chorus is a
public group Facebook published notifications to Duncan's and McCormick's
friends, including their parents. Duncan later attempted suicide~\cite{duncan}.

@@

Barth, Datta, Mitchell, and Nissenbaum provide a formal model for contextual
integrity of user data~\cite{barth}. The contextual integrity model focuses on
the flow of personal information in terms of knowledge states held by agents
acting in given roles, and works well for modeling privacy legislation such as
HIPPA and COPPA. Two drawbacks of this work are that it does not
model information where the subject involves more than one person, and it
only takes the type of information into account rather than the content or tone
of the information.  ``Bob told Alice that Charlie cooked a
delicious dinner'' is out of scope of the model.  In the context of social
networks, information is rarely about a single person, and the tone of the
message is as important to user effect as the content of the message.

@@

\section{Contextual identity violations}
We consider it a contextual identity violation when
multiple contextual identities are linked without the user's permission, or the
user cannot control which contextual identity is presented to a given audience.

\label{sec:examples}
\subsection{Redistributing information out of context}
Re-distributing information out of its original context often leads to
embarrassment~\cite{nissenbaum}.  In November 2007, Facebook Beacon allowed
third-party sites to publish purchases, travel bookings, movie rentals and more
to the user's activity stream.  Because of poor opt-out and lack of visibility
into what was being published, user outcry was
immediate~\cite{mccarthy,nakashima}.  In December 2007, Google Reader exposed
RSS feeds of user-marked news stories to the user's Google Talk contacts, which
includes everyone with whom a user has chatted whether they be co-workers,
supervisors, or friends. Although this feed was always public, prior to this
launch it was not discoverable, leading to a bad experience for many
users~\cite{helft}. In February 2010, Google Buzz launched, a product that
exposed the user's most frequent Gmail and Google Talk contacts and their
publicly available (through previously less discoverable) news and photos.  The
combination of exposing contacts (which could include abusive ex-husbands,
co-workers, and friends), as well as aggressively linking photos and news
items, resulted in a huge backlash~\cite{fugitivus,buzz}. In September 2012,
Facebook imported old wall posts from 2008 into the new Timeline interface.
Although wall posts were always visible from profile pages, the new Timeline
interface brought old wall posts which users used to treat as private messages,
before the advent of ``Like'' and comment buttons, to the attention of an
audience that the user never anticipated~\cite{timeline}.

\subsection{Policy errors}
\label{sec:policies}
Some service providers have policies which preclude isolating multiple
contextual identities. For example, Facebook and Google have a ``Real Names''
policy, which requires users to register for accounts with their legal
name~\cite{fb_names,google_names}.  These policies ignore that
community-building happens in many different contexts, that users have many
legitimate reasons for presenting different identities in different contexts,
and that users don't necessarily want those identities to be linked. For
example, disallowing avatar handles as a primary identifier makes building a
gaming community difficult.  Linking a gaming identity to a professional
identity has already caused problems for 2012 Maine Senate candidate and World
of Warcraft gamer Colleen Lachowicz~\cite{maine}.

@@

\section{Research directions}
We propose new research directions based on the following questions.
\begin{itemize}
\item How do users think about identity?
\item What can we infer about identity?
\item What threat models (audiences, topics) are important to users?
\item What are users already doing to preserve contextual identities?
\end{itemize}

\subsection{How do users think about identity?}
Users' contextual identities may be task-oriented, service-oriented,
relationship-oriented, or guided by other principles. The long-time existence
of support for multiple accounts in email clients suggests that many users have
multiple email identifiers, which could be considered as a proxy for identity
if tied to service accounts. Similarly, data suggests that a large minority of
Twitter users have multiple Twitter accounts, and the plethora of cross-posting
or syndication software from one social network to another suggests that many
users take advantage of multiple social networks~\cite{twitter,xpost}.  That is
not to say that all users have the same tolerance or even need for multiple
explicit identities: in fact many users may be exclusively task-oriented and
use the Internet with specific tasks in mind rather than for social interaction
or exploratory surfing\footnote{Even these users are subject to building
implicit identities through the use of tracking techniques, however.}.

Although a user may subconsciously be aware of their contextual identities,
expecting users to enumerate their identities and manually curate them is
probably too burdensome. Firefox has a Profile Manager, but it is so difficult
to use that it will be removed.\footnote{See
\url{https://bugzilla.mozilla.org/show\_bug.cgi?id=214675\#c53}.} Tang et
al.~show that most people do not take the time to create labels and maintain
friend lists~\cite{tang}. However, the same authors show how to use
heterogeneous networks to infer relationship types (e.g., manager-subordinate)
though the graph itself is unlabeled.

Further research is required to determine what constitutes a contextual
identity. The contextual integrity model from Barth et al.~does not capture
these notions adequately.

(This is weak)
If users' contextual identities are primarily task oriented, it may be
possible to infer tasks from the user's activity. For example, the
\texttt{about:profile} Firefox extension uses the Open Directory Project and
Alexa to categorize URLs that the user has visited~\cite{aboutprofile}.
